---
title: "Stage 1 Growth / Cage Data Load"
author:
- name: Robert Sellers
  affiliation: The Jackson Laboratory
date: "`r paste('Last knit on:',format(Sys.time(), '%d %B %Y'))`"
abstract: |
latex_engine: pdflatex
output:
  html_document:
    fig_caption: yes
    force_captions: yes
    number_sections: true
    theme: lumen
    toc: yes
    toc_float: yes
    df_print: paged
    # code_folding: "hide"
  pdf_document:
    fig_caption: yes
    force_captions: yes
    highlight: pyments
    keep_tex: yes
    latex_engine: xelatex
    number_sections: no
    toc: no
    
---

# Environment

```{r}
options(scipen = 999) # suppress sci notation

setwd('~/Desktop/projects/jax_arsenic_toxicology_analysis')

suppressPackageStartupMessages(require(tidyverse))
suppressPackageStartupMessages(require(readxl))
suppressPackageStartupMessages(require(ggpubr))
suppressPackageStartupMessages(require(zoo))

source('scripts/custom_tools.R')
```

# Data & custom changes

## Supplementary cage data

```{r}
housing <- read_excel(
    "./data/invivo/arsenic_mouse_bodyweight.xlsx",
    sheet = "housing",
    .name_repair = janitor::make_clean_names
  ) %>%
  separate(name, c("housing_var_a","id","housing_var_c"), sep='-') %>%
  select(-one_of("housing_var_a","housing_var_c")) 
```

## Raw bodyweights data

```{r}
bodyweights <- read_excel(
    "./data/invivo/arsenic_mouse_bodyweight.xlsx",
    sheet = "bodyweight",
    .name_repair = janitor::make_clean_names
  ) %>%
  dplyr::rename(generation = g) %>%
  mutate(
    id = factor(id),
    date_of_birth = as.Date(dob,"%m/%d/%y", tz = "UTC")
  ) %>%
  select(-dob) %>%
  left_join(housing) %>%
  group_by(housing_id) %>%
  arrange(id) %>%
  # adds a unique color column per id found
  group_modify(~ {
    .x %>%
      mutate(color_gen = sample(scales::hue_pal()(nrow(.))))
  }) %>%
  gather(week, bodyweight, bw8:bw52) %>%
  mutate(
   # extract numeric week integer
   weeks_continuous = readr::parse_number(week),
   # adds week integer to real date
   week_as_date = date_of_birth + lubridate::weeks(weeks_continuous),
   # round bodyweight to one decimal, NA warning is fine
   bodyweight = suppressWarnings(round(as.numeric(bodyweight),2))) %>%
  # handle duplicates
  group_by(id, weeks_continuous) %>% 
  mutate(num_dups = n(), 
         dup_id = row_number()) %>% 
  ungroup() %>% 
  mutate(id = case_when(num_dups > 1 ~ paste0("D",dup_id,"_",id),
         TRUE ~ id)) %>%
  mutate(fdose = factor(dose)) %>% 
  select(-one_of("week","dup_id","num_dups","dose")) %>%
  # add date ranking for normal date analysis
  group_by(wave, id) %>%
  mutate(date_order = order(order(week_as_date))) %>%
  #############
  # manual fixes
 # incorrect names or doses / modifies duplicates
  mutate(id=ifelse((id=='D2_10556'), '10556', id),
         id=ifelse((id=='D1_10658'),'10656', id),
         id=ifelse((id=='D2_10658'),'10658', id)) %>%
  # add_column(notes = NA) %>%
  mutate_cond(housing_id == 'H-89-19', fdose = factor(1)) %>%
  mutate_cond(housing_id == 'H-85-19', fdose = factor(0.1)) %>%
  # measurement errors
  # these data will get picked up for mean interpolation
  # do not attempt on first or last values in series
  mutate_cond(id =='10464' & date_order == 11, bodyweight = NA) %>%
  mutate_cond(id =='10261' & date_order == 8, bodyweight = NA) %>%
  mutate_cond(id =='10664' & date_order == 8, bodyweight = NA) %>%
  mutate_cond(id =='10189' & date_order == 10, bodyweight = NA) %>%
  mutate_cond(id =='10190' & date_order == 10, bodyweight = NA) %>%
  mutate_cond(id =='10309' & date_order == 8, bodyweight = NA) %>%
  ### annotate changes
  mutate(notes = case_when(
      (id == "10464" & date_order == 11) | (id =='10261' & date_order == 8) |
      (id =='10664' & date_order == 8) | (id =='10189' & date_order == 10) |
      (id =='10190' & date_order == 10) | (id =='10309' & date_order == 8) 
      ~ "mean_interpolated", TRUE   ~ "original" ))
```

## Documented deaths data

```{r}
# provided via email 
deaths_supplement <- read_excel(
    "./data/invivo/arsenic_mouse_bodyweight.xlsx",
    sheet = "deaths_supplement",
    .name_repair = janitor::make_clean_names
  )  %>%
  mutate(
    death_exit_date = as.Date(death_exit_date, tz = "UTC"),
    id = as.character(id)
  ) %>%
  dplyr::rename(week_as_date = death_exit_date) %>%
  filter(id %notin% c('100058','100059','10194',"100033","10391","10513","10515",  "10652","10686","10219","10316","10339")) %>%
  # need to fill dob 
  left_join(dplyr::distinct(bodyweights[,c("id","date_of_birth")],id,.keep_all = T)) %>%
  mutate(weeks_continuous =as.numeric((week_as_date - date_of_birth)/7)) %>%
  dplyr::select(-date_of_birth)

deaths <- read_excel( 
    "./data/invivo/arsenic_mouse_bodyweight.xlsx",
    sheet = "deaths",
    .name_repair = janitor::make_clean_names
  )  %>%
  mutate(
    death_exit_date = as.Date(death_exit_date, tz = "UTC"),
    cause_of_death = status
  ) %>%
  dplyr::rename(week_as_date = death_exit_date,
                weeks_continuous = age_weeks) %>%
  separate(name, c("housing_var_a","id","housing_var_c"), sep='-') %>%
  select(-one_of("housing_var_a","housing_var_c", "status")) %>%
  left_join(housing)  %>%
  select(-one_of("sex","birth_date")) %>%
  dplyr::bind_rows(deaths_supplement) %>%
  mutate(
    data_source = 'documented'
    ) %>%
  plyr::join(dplyr::distinct(bodyweights[,c('fdose','color_gen','id','fdose','sex','wave','generation')],id,.keep_all = T), by = 'id')

# creates dataset from final non-NA value encountered or documented deaths
estimated_deaths <- bodyweights %>%
  mutate(
    data_source = "end_of_study"
    ) %>%
  drop_na(bodyweight) %>%
  bind_rows(deaths) %>%
  group_by(id) %>%
  filter(week_as_date == max(week_as_date)) %>%
  select(-date_order,date_of_birth) %>%
  group_by(id) %>%
  # filter out instances where documented and end_of_study share the same date
  # documented before end_of_study
  arrange(data_source, .by_group = TRUE) %>% 
  filter(row_number() == 1)
```

# Predictive analysis 

## Fit logic / data smoothing

- uses a span 'gradient' for # of points
- uses lm on 1 or 2 data points, with loess on 3+

```{r}
try_loess_lm <- function (df) {
  return(
    tryCatch({
      if (nrow(na.omit(df)) == 2 | nrow(na.omit(df)) == 1){
        lm(bodyweight ~ weeks_continuous, data = na.omit(df))
      }else if (nrow(na.omit(df)) == 3){
        loess(bodyweight ~ weeks_continuous, data = na.omit(df),control = loess.control(surface="direct"), span = 2)
      }else if(nrow(na.omit(df)) > 20){
        loess(bodyweight ~ weeks_continuous, data = na.omit(df), span = 0.25)
      }else{
        # default
        loess(bodyweight ~ weeks_continuous, data = na.omit(df),control = loess.control(surface="direct"), span = 1)
      }
    }, error=function(e) {
        paste0("failed with ", nrow(na.omit(df)))
      }
    ))
}

try_predict <- function (x, df){
    return(tryCatch(suppressWarnings(predict(x, df)), error=function(e) NULL))
}

# generate loess/lm models
loess_lm_models <- plyr::dlply(bodyweights, "id", function(df){
  suppressWarnings(try_loess_lm(df))
})
# predict data where is documented death recorded
# WARNING: this is not finding duplicated models, but this wasn't necessary here for these 3/6 dups
# many model are returning as NULL
sub_models <- loess_lm_models[unlist(names(loess_lm_models) %in% estimated_deaths[estimated_deaths$data_source == 'documented', ]$id)]

# extracts predictive inputs from loess model list
sub_dates <- subset(estimated_deaths, id %in% names(sub_models))[,c('id','weeks_continuous')] %>%
  arrange(id)

# run predictions
predictions <- list()
for (i in seq_along(sub_models)) {
  pred_name <- names(sub_models)[i]
  # implement better error catching here
  predictions[pred_name] <- try_predict(
    x = sub_models[[i]], 
    df = data.frame("weeks_continuous"= sub_dates$weeks_continuous[i])
  )
}

# cleanup
pred_df <- stack(predictions) %>%
   dplyr::rename(id = ind, bodyweight_fit = values)

# apply predictions to back dataframe and annotate changes
cause_of_death <- estimated_deaths %>%
  plyr::join(pred_df, by = 'id') %>%
  mutate(week_as_date = as.Date(week_as_date, tz = "UTC")) %>%
  group_by(wave) %>%
  arrange(date_of_birth) %>%
  mutate(date_of_birth = date_of_birth[1]) %>%
  mutate_cond(!is.na(bodyweight_fit), notes = 'estimated') %>%
  mutate_cond(data_source == 'documented' & is.na(bodyweight_fit), notes = 'no_data')
```

## Impute / interpolation

- operates on NA between existing values

```{r}
bodyweights <-  bodyweights %>%
  group_by(id) %>%
  mutate(bodyweight_test = na.approx(bodyweight, na.rm=FALSE)) %>%
  mutate_cond(is.na(bodyweight) & !is.na(bodyweight_test), bodyweight = bodyweight_test) %>%
  dplyr::select(-bodyweight_test)
```

# Model fitting / data smoothing

- applies available models to each mouse's raw data as is filtered inside try_loess_lm()
- combine with death estimates

```{r}
bw_fit_w_estimates <- plyr::ldply(loess_lm_models, function(mod){
  # first construct a data.frame for entire list of models x / y
  if (class(mod) == "loess"){
        data.frame("weeks_continuous" = mod$x, "bodyweight" = mod$fitted)
      }
    }) %>%
  mutate(tmp_id = gsub(".*_", "", id)) %>%
  left_join(housing, by = c("tmp_id" = "id")) %>%
  right_join(bodyweights, 
             by=c("id","weeks_continuous"),
             suffix = c("_fit", "_recorded")) %>%
  select(-one_of("tmp_id","housing_id_fit")) %>% 
  dplyr::rename(housing_id = housing_id_recorded,
                bodyweight = bodyweight_recorded) %>%
  bind_rows(cause_of_death)
```

# Saves

```{r}
write.csv(bw_fit_w_estimates, './output/bodyweight/bw_processed.csv', row.names = FALSE)
```