---
title: "Stage 1 Growth / Cage Data Load"
author:
- name: Robert Sellers
  affiliation: The Jackson Laboratory
date: "`r paste('Last knit on:',format(Sys.time(), '%d %B %Y'))`"
abstract: |
latex_engine: pdflatex
output:
  html_document:
    fig_caption: yes
    force_captions: yes
    number_sections: true
    theme: lumen
    toc: yes
    toc_float: yes
    df_print: paged
    # code_folding: "hide"
  pdf_document:
    fig_caption: yes
    force_captions: yes
    highlight: pyments
    keep_tex: yes
    latex_engine: xelatex
    number_sections: no
    toc: no
    
---

# Environment

```{r}
options(scipen = 999) # suppress sci notation

setwd('~/Desktop/projects/jax_arsenic_toxicology_analysis')

suppressPackageStartupMessages(require(tidyverse))
suppressPackageStartupMessages(require(readxl))
suppressPackageStartupMessages(require(ggpubr))
suppressPackageStartupMessages(require(zoo))

source('scripts/custom_tools.R')
```

# Data & custom changes

## Supplementary cage data

```{r}
housing <- read_excel(
    "./data/invivo/arsenic_bodyweight.xlsx",
    sheet = "housing",
    .name_repair = janitor::make_clean_names
  ) %>%
  separate(name, c("housing_var_a","id","housing_var_c"), sep='-') %>%
  select(-one_of("housing_var_a","housing_var_c")) 
```

## Raw bodyweights data

```{r}
bodyweights_raw <- read_excel(
    "./data/invivo/arsenic_bodyweight.xlsx",
    sheet = "bodyweight",
    .name_repair = janitor::make_clean_names
  ) 

bodyweights <- bodyweights_raw %>%
  dplyr::rename(generation = g) %>%
  mutate(
    id = factor(id),
    date_of_birth = as.Date(dob,"%m/%d/%y", tz = "UTC")
  ) %>%
  select(-dob) %>%
  left_join(housing) %>%
  group_by(housing_id) %>%
  arrange(id) %>%
  # adds a unique color column per id found
  group_modify(~ {
    .x %>%
      mutate(color_gen = sample(scales::hue_pal()(nrow(.))))
  }) %>%
  gather(week, bodyweight, bw8:bw52) %>%
  mutate(
   # extract numeric week integer
   weeks_continuous = readr::parse_number(week),
   # adds week integer to real date
   week_as_date = date_of_birth + lubridate::weeks(weeks_continuous),
   # round bodyweight to one decimal, NA warning is fine
   bodyweight = suppressWarnings(round(as.numeric(bodyweight),2))) %>%
  # handle duplicates
  group_by(id, weeks_continuous) %>% 
  mutate(num_dups = n(), 
         dup_id = row_number()) %>% 
  ungroup() %>% 
  mutate(id = case_when(num_dups > 1 ~ paste0("D",dup_id,"_",id),
         TRUE ~ id)) %>%
  mutate(fdose = factor(dose)) %>% 
  select(-one_of("week","dup_id","num_dups","dose")) %>%
  # add date ranking for normal date analysis
  group_by(wave, id) %>%
  mutate(date_order = order(order(week_as_date))) %>%
  #############
  # manual fixes
 # incorrect names or doses / modifies duplicates
  mutate(id=ifelse((id=='D2_10556'), '10556', id),
         id=ifelse((id=='D1_10658'),'10656', id),
         id=ifelse((id=='D2_10658'),'10658', id)) %>%
  # add_column(notes = NA) %>%
  mutate_cond(housing_id == 'H-89-19', fdose = factor(1)) %>%
  mutate_cond(housing_id == 'H-85-19', fdose = factor(0.1)) %>%
  # measurement errors
  # these data will get picked up for mean interpolation
  # do not attempt on first or last values in series
  mutate_cond(id =='10464' & date_order == 11, bodyweight = NA) %>%
  mutate_cond(id =='10261' & date_order == 8, bodyweight = NA) %>%
  mutate_cond(id =='10664' & date_order == 8, bodyweight = NA) %>%
  mutate_cond(id =='10189' & date_order == 10, bodyweight = NA) %>%
  mutate_cond(id =='10190' & date_order == 10, bodyweight = NA) %>%
  mutate_cond(id =='10309' & date_order == 8, bodyweight = NA) %>%
  ### annotate changes
  mutate(notes = case_when(
      (id == "10464" & date_order == 11) | (id =='10261' & date_order == 8) |
      (id =='10664' & date_order == 8) | (id =='10189' & date_order == 10) |
      (id =='10190' & date_order == 10) | (id =='10309' & date_order == 8) 
      ~ "mean_interpolated", TRUE   ~ "original" ))
```

## Documented deaths data

```{r}
# provided via email 
deaths_supplement <- read_excel(
    "./data/invivo/arsenic_bodyweight.xlsx",
    sheet = "deaths_supplement",
    .name_repair = janitor::make_clean_names
  )  %>%
  mutate(
    death_exit_date = as.Date(death_exit_date, tz = "UTC"),
    id = as.character(id)
  ) %>%
  dplyr::rename(week_as_date = death_exit_date) %>%
  filter(id %notin% c('100058','100059','10194',"100033","10391","10513","10515",  "10652","10686","10219","10316","10339")) %>%
  # need to fill dob 
  left_join(dplyr::distinct(bodyweights[,c("id","date_of_birth")],id,.keep_all = T)) %>%
  mutate(weeks_continuous =as.numeric((week_as_date - date_of_birth)/7)) %>%
  dplyr::select(-date_of_birth)

deaths <- read_excel( 
    "./data/invivo/arsenic_bodyweight.xlsx",
    sheet = "deaths",
    .name_repair = janitor::make_clean_names
  )  %>%
  mutate(
    death_exit_date = as.Date(death_exit_date, tz = "UTC"),
    cause_of_death = status
  ) %>%
  dplyr::rename(week_as_date = death_exit_date,
                weeks_continuous = age_weeks) %>%
  separate(name, c("housing_var_a","id","housing_var_c"), sep='-') %>%
  select(-one_of("housing_var_a","housing_var_c", "status")) %>%
  left_join(housing)  %>%
  select(-one_of("sex","birth_date")) %>%
  dplyr::bind_rows(deaths_supplement) %>%
  mutate(
    data_source = 'documented'
    ) %>%
  plyr::join(dplyr::distinct(bodyweights[,c('fdose','color_gen','id','fdose','sex','wave','generation')],id,.keep_all = T), by = 'id')

# creates dataset from final non-NA value encountered or documented deaths
estimated_deaths <- bodyweights %>%
  mutate(
    data_source = "end_of_study"
    ) %>%
  drop_na(bodyweight) %>%
  bind_rows(deaths) %>%
  group_by(id) %>%
  filter(week_as_date == max(week_as_date)) %>%
  select(-date_order,date_of_birth) %>%
  group_by(id) %>%
  # filter out instances where documented and end_of_study share the same date
  # documented before end_of_study
  arrange(data_source, .by_group = TRUE) %>% 
  filter(row_number() == 1)
```

# Predictive analysis 

## Fit logic / data smoothing

- uses a span 'gradient' for # of points
- uses lm on 1 or 2 data points, with loess on 3+

```{r}
# generate loess/lm models
loess_lm_models <- plyr::dlply(bodyweights, "id", function(df){
  suppressWarnings(try_loess_lm(df))
})
# predict data where is documented death recorded
# WARNING: this is not finding duplicated models, but this wasn't necessary here for these 3/6 dups
# many model are returning as NULL
sub_models <- loess_lm_models[unlist(names(loess_lm_models) %in% estimated_deaths[estimated_deaths$data_source == 'documented', ]$id)]

# extracts predictive inputs from loess model list
sub_dates <- subset(estimated_deaths, id %in% names(sub_models))[,c('id','weeks_continuous')] %>%
  arrange(id)

# run predictions
predictions <- list()
for (i in seq_along(sub_models)) {
  pred_name <- names(sub_models)[i]
  # implement better error catching here
  predictions[pred_name] <- try_predict(
    x = sub_models[[i]], 
    df = data.frame("weeks_continuous"= sub_dates$weeks_continuous[i])
  )
}

# cleanup
pred_df <- stack(predictions) %>%
   dplyr::rename(id = ind, bodyweight_fit = values)

# apply predictions to back dataframe and annotate changes
cause_of_death <- estimated_deaths %>%
  plyr::join(pred_df, by = 'id') %>%
  mutate(week_as_date = as.Date(week_as_date, tz = "UTC")) %>%
  group_by(wave) %>%
  arrange(date_of_birth) %>%
  mutate(date_of_birth = date_of_birth[1]) %>%
  mutate_cond(!is.na(bodyweight_fit), notes = 'estimated') %>%
  mutate_cond(data_source == 'documented' & is.na(bodyweight_fit), notes = 'no_data')
```

## Impute / interpolation

- operates on NA between existing values
- this is extremely limited - needs a lot of work

```{r}
bodyweights <-  bodyweights %>%
  group_by(id) %>%
  mutate(bodyweight_test = na.approx(bodyweight, na.rm=FALSE)) %>%
  mutate_cond(is.na(bodyweight) & !is.na(bodyweight_test), bodyweight = bodyweight_test) %>%
  dplyr::select(-bodyweight_test)
```

# Model fitting / data smoothing

- applies available models to each mouse's raw data as is filtered inside try_loess_lm()
- combine with death estimates

```{r}
bw_fit_w_estimates <- plyr::ldply(loess_lm_models, function(mod){
  # first construct a data.frame for entire list of models x / y
  if (class(mod) == "loess"){
        data.frame("weeks_continuous" = mod$x, "bodyweight" = mod$fitted)
      }
    }) %>%
  mutate(tmp_id = gsub(".*_", "", id)) %>%
  left_join(housing, by = c("tmp_id" = "id")) %>%
  right_join(bodyweights, 
             by=c("id","weeks_continuous"),
             suffix = c("_fit", "_recorded")) %>%
  select(-one_of("tmp_id","housing_id_fit")) %>% 
  dplyr::rename(housing_id = housing_id_recorded,
                bodyweight = bodyweight_recorded) %>%
  bind_rows(cause_of_death) %>%
  mutate(bodyweight_fit= round(bodyweight_fit,2)) %>%
  relocate(id, bodyweight, bodyweight_fit,date_order,sex, housing_id)
```

# Annotated dataset construction

- complementary / 1:many table for analysis
- https://stackoverflow.com/questions/12851208/how-to-detect-significant-change-trend-in-a-time-series-data

```{r}
bw_annotated <- read.csv("./data/invivo/arsenic_bodyweight_annotated.csv")
```

# Saves

## Bodyweight time series

- currently filters 1) redundant 'end_of_study' extraction and 2) empty gathered fields.

```{r}
write.csv(dplyr::select(filter(bw_fit_w_estimates, !is.na(date_order) & !is.na(bodyweight)), -c("color_gen", "data_source","comments","cause_of_death")), './output/bodyweight/arsenic_bodyweight_processed.csv', row.names = FALSE)
```

## Fits plot

 - not knit - look for PDF

```{r}
# list no data
na_list <- bw_fit_w_estimates %>%
  group_by(id,housing_id,wave,generation,fdose, sex) %>% 
  summarise_each(funs(100*mean(is.na(.)))) %>%
  filter(bodyweight == 100) %>%
  group_by(housing_id,sex,wave,fdose,generation) %>%
  summarise(label_nodata = paste("no data:", id, collapse = "\n"),
            label_group = paste(paste0("wave:",wave), housing_id, paste0("wave:",sex), paste0("wave:",fdose), collapse = "\n"))
```

```{r eval = FALSE}
pdf("./output/bodyweight/mice_by_cage.pdf")
# use n_pages on ggplot object on one page to estimate loop size
#ggforce::n_pages()

for (i in 1:ceiling(length(unique(bw_fit_w_estimates$housing_id))/5)){
  print(paste0("producing  plot page #", i))
  print(
  ggplot(arrange(bw_fit_w_estimates, sex, wave, housing_id, fdose), 
    aes(x = week_as_date, y = bodyweight, group = id)) +
  geom_path(aes(color = color_gen), 
            size = 0.5,
            linetype = 'dotdash') +
      geom_text(data = na_list, x=-Inf, y=Inf, aes(group = housing_id,label = label_nodata),hjust = 0, vjust = 1, size=3) +
  scale_shape_manual(values = c(16, 1)) +   
    # smooths on lm results
  geom_smooth(data = filter(bw_fit_w_estimates,id %in% names(which(sapply(loess_lm_models, function(x) (class(x) == "lm")),TRUE))) %>% mutate_cond(is.na(bodyweight), bodyweight = bodyweight_fit),
              method = 'lm',
              aes(x = week_as_date,
                  y = bodyweight,
                  color = color_gen),
              linetype = 'solid',
              se = FALSE,
              size=0.5) +
    # smooths on loess results
  geom_smooth(data = filter(bw_fit_w_estimates,id %in% names(which(sapply(loess_lm_models, function(x) (class(x) == "loess")),TRUE)))  %>% mutate_cond(is.na(bodyweight), bodyweight = bodyweight_fit),
              method = 'loess',
              aes(x = week_as_date,
                  y = bodyweight,
                  color = color_gen),
              linetype = 'solid',
              se = FALSE,
              span = 1,
              size = 0.5) +
  geom_point(data = cause_of_death %>% mutate_cond(is.na(bodyweight), bodyweight = bodyweight_fit),
             size = 2, show.legend = FALSE,
             aes(x = week_as_date, 
                 y = bodyweight,
                 shape = data_source,
                 color = color_gen)
             ) +
  ggrepel::geom_text_repel(
    aes_q(label = ~paste(id,"\n", 
          ifelse(is.na(cause_of_death),"",paste0(sprintf("%s",cause_of_death),"\n")),
          ifelse(is.na(comments),"",sprintf("%s",comments))), 
           color = as.character(cause_of_death$color_gen)),
           fontface = "bold",
           direction = 'both',
           nudge_x = 25,
           bg.color = "white",
           bg.r = 0.2,
           force_pull = 3,
           segment.colour = 'black',
           min.segment.length = unit(0, 'lines'),
           data =  cause_of_death %>% mutate_cond(is.na(bodyweight), bodyweight = bodyweight_fit),
           size = 2)  + 
          geom_point(data = bw_fit_w_estimates, aes(x = week_as_date, y = bodyweight), size=0.15) +
  ######### setting x limits, nudging to the right #########
ggforce::facet_wrap_paginate(~ housing_id + sex + fdose + wave, 
                               ncol = 1, 
                               nrow = 5,
                               page  = i,
                               strip.position="right",
                               scales = "free_y",
                               labeller = labeller( wave = label_both, housing_id = label_both, fdose = label_both, sex = label_both)) + 
  scale_x_date(date_labels = "%b-%y", date_breaks = '1 month',minor_breaks=NULL) +
  ylab("Bodyweights (grams)") +
  xlab("Arsenic Dose (ppm)") +
  theme(
    strip.text = element_text(size=8),
    panel.border = element_rect(fill=NA, colour="grey40"),
    axis.text.x = element_text(size = 6, angle = 45, vjust = 1.1, hjust=1),
    axis.title.x=element_blank()
  ) + guides(colour = FALSE, shape = guide_legend(),
  linetype = guide_legend())
  )
}
dev.off()
```