---
title: "Founders #1 Preprocess and QA"
author:
- name: Robert Sellers
  affiliation: The Jackson Laboratory
date: "`r paste('Last knit on:',format(Sys.time(), '%d %B %Y'))`"
abstract: |
latex_engine: pdflatex
output:
  html_document:
    fig_caption: yes
    force_captions: yes
    number_sections: true
    theme: lumen
    toc: yes
    toc_float: yes
    df_print: paged
    # code_folding: "hide"
  pdf_document:
    fig_caption: yes
    force_captions: yes
    highlight: pyments
    keep_tex: yes
    latex_engine: xelatex
    number_sections: no
    toc: no
    
---

# Programming environment

```{r message=FALSE}
# contains harmony plate collection class builder
source("scripts/harmony_utils_v2.R")
# various add-on scripts and stats tools
source("scripts/custom_tools.R")

library(tidyverse)
library(janitor)
library(ggplot2) 
library(ggpubr) # ggplot extension
```

# Load data

## Build plate collection from directory

```{r warning=FALSE, echo = FALSE}
# point to parent data directory
setwd("data/discovery_invitro/")
selected_dir <- "./founders_quantal"
# Warning - this currently expects a temp_files data dir 
# Set overwrite to TRUE when underlying data dir has been changed
founders_quantal_collection <- harmony_create_collection(dir = selected_dir, overwrite = TRUE)
#do_focus_collection <- harmony_create_collection(dir = './do_focus',overwrite = TRUE) 
```

## Data treatments

```{r warning = FALSE}
# consult https://cran.r-project.org/web/packages/janitor/vignettes/janitor.html
founders_quantal_collection$all_plates <- founders_quantal_collection$all_plates %>% 
  dplyr::rename(
    strain = cell_type, 
    dose = concentration
    ) %>%
  mutate(
    plate = case_when(
           dir=="dFounders_A1_Whole_New__20200115T13_25_19Measurement_1_e1" ~ "A1",
           dir=="dFounders_B1_Whole__20200103T10_34_56Measurement_1_e3" ~ "B1",
           dir=="dFounders_AA1__20200113T14_27_40Measurement_1_e3" ~ "AA1",
           dir=="dFounders_BB1__20200114T09_14_08Measurement_1_e3" ~ "BB1",
           dir=="dFounders_CC1__20200115T09_21_08Measurement_1_e3" ~ "CC1",
           dir=="dFounders_D1_Whole__20191230T13_52_41Measurement_1_e2" ~ "D1",
           dir=="dFounders_DD1__20200114T13_08_38Measurement_1_e3" ~ "DD1",
           dir=="dFounders_E1_Whole__20200102T08_52_12Measurement_1_e4" ~ "E1",
           dir=="dFounders_C1_Whole__20191230T10_37_26Measurement_1_e2" ~ "C1"
         )
    ) %>%
  filter((plate == 'B1' & mouse != "PWK_2F") | plate != 'B1') %>%
  # remove duplicates
  filter(mouse %notin% c("AJ_2F", "B6_2F")) %>%
  mutate(sex = case_when(plate == 'D1' ~ "F",TRUE ~ as.character(sex))) %>%
  # need to confirm this again on harmony
  mutate(strain = if_else(is.na(strain) & plate == 'A1', "CAST", strain)) %>%
  mutate(
    strain = factor(strain),
    sex = factor(sex),
    dose = factor(dose))  %>% 
  mutate(mouse = case_when(
         mouse == 'CAST_2M' ~ "CAST_M2",
         mouse == 'CAST_2F' ~ 'CAST_F2',
         mouse == 'AJ_2M' ~ 'AJ_M2',
         mouse == 'AJ_2F' ~ 'AJ_F2',
         mouse == 'B6_2M' ~ 'B6_M2',
         mouse == 'PWK_2F' ~ 'PWK_F2',
         mouse == '129_2F' ~ '129_F2',
         mouse == 'NOD_2M' ~ 'NOD_M2',
         mouse == 'NOD_2F' ~ 'NOD_F2',
         mouse == 'PWK_2M' ~ 'PWK_M2',
         mouse == 'B6_2F' ~ 'B6_F2',
         TRUE ~ mouse)
     ) %>%
  dplyr::select(plate,row,column,dose,strain,sex,mouse, everything())
```

## Coordinated feature metadata
- associates annotated info (e.g. stain)
- removes data not represented in both founders + DO based on vlookup in google sheet

```{r}
# pseudo database solution
# you can't do this too frequently < every 2 minutes or so
feature_metadata <- googlesheets4::read_sheet(ss = "1A6qMVGhtJykfbCQK1qq0qv5w3HfKvs8ByVSCiExc49s", sheet = 'founders_quantal') 

# remove unmatched features from all_plates
# this doesn't seem necessary for Founders but will need for DO
founders_quantal_collection$all_plates <- founders_quantal_collection$all_plates %>% select (-feature_metadata$name_[feature_metadata$mutuality==FALSE])

# remove unmatched features from features metadata (if applicable)
founders_quantal_collection$features_df <- founders_quantal_collection$features_df %>%
  left_join(feature_metadata[,c("name_","stain","mutuality")], "name_") %>%
  filter(mutuality) %>% # TRUE/FALSE if in DO and vice versa
  select(-mutuality)
```

## Remove STDEV data
- Targets both all_plates & feature_df currently

```{r}
# grep function that removes a pattern
remove_sd_vars <- function(data){
  # remove from all_plates
  before_cols <- ncol(data$all_plates)
  data$all_plates <- dplyr::select(data$all_plates, -dplyr::contains("stdev"))
  print(paste("removing",(before_cols-ncol(data$all_plates)), "/", before_cols,"features"))
  # remove from features_df
  data$features_df <- data$features_df[!grepl("stdev", data$features_df$name_),]
  return (data)
}

# remove columns / features with pattern "stdev"
founders_quantal_collection <- remove_sd_vars(founders_quantal_collection)
```

# Save/overwrite RDS

```{r}
temp_data_storage <- paste0("temp_files/",selected_dir,'_processed.RData')
saveRDS(founders_quantal_collection, temp_data_storage)
```

---

- __Continue this pipeline with founders_data_transform.rmd and while using the _processed.RData_ dataset__