---
title: "Stage 1 DO Harmony Initial Data Load"
author:
- name: Robert Sellers
  affiliation: The Jackson Laboratory
date: "`r paste('Last knit on:',format(Sys.time(), '%d %B %Y'))`"
abstract: |
latex_engine: pdflatex
output:
  html_document:
    fig_caption: yes
    force_captions: yes
    number_sections: true
    theme: lumen
    toc: yes
    toc_float: yes
    df_print: paged
    # code_folding: "hide"
  pdf_document:
    fig_caption: yes
    force_captions: yes
    highlight: pyments
    keep_tex: yes
    latex_engine: xelatex
    number_sections: no
    toc: no
    
---

Document/Process explained:

* The following loads a "raw" bound object of all 96 well plates for an experiment. This data is available elsewhere as _do_focus_merge_2_ for raw and metadata .csv / .tsv files.

* The resulting object list contains both the loaded data __all_plates__ with __all_plates_clean__ data.frames.

* A subset of curated features is also constructed / referenced as 'harmony_do_focus_reduced.csv'.

* these objects can be simply loaded as follows: 
  
```{r eval = FALSE}
harmony_collection <- readRDS('./harmony_do_focus.RData')
analytics_csv <- read.csv('./harmony_do_focus_reduced.csv')
```

# Environment

- Point your working directory to the project root folder

```{r message=FALSE}
# env
setwd('~/Desktop/projects/jax_arsenic_toxicology_analysis/')
options(scipen = 999) # suppress sci notation
# contains harmony plate collection class builder
source("scripts/harmony_utils.R")
# warehouse for functions
source("scripts/custom_tools.R") 
require(dplyr)
```

# Data

## Direct load from .rdata or .csv

- Use this if not working with Harmony software data or development.

```{r}
harmony_collection <- readRDS('./harmony_do_focus.RData')
analytics_csv <- read.csv('./harmony_do_focus_reduced.csv')
```

## Harmony_collection build (if applicable)

__The next two code blocks are optional / not knitted. This utility is a development tool.__

### Build collection from raw option

- The following code block is set to eval=FALSE
- overwrite = TRUE builds a temp .rds file (e.g. temp_files/do_focus_merge_2.rds) from raw harmony data.

```{r warning=FALSE, eval=FALSE}
# point to parent data directory
selected_dir <- "data/invitro/harmony_do_focus"
# Warning - this logic expects a ./temp_files data dir for caching
# Set overwrite to TRUE when underlying data dir has been changed
harmony_collection <- harmony_create_collection(dir = selected_dir,
                                                overwrite = TRUE,
                                                temp_dir = "./temp_files",
                                                custom_rename = FALSE)
```

### Load collection from cache

- The following code block is set to eval=FALSE
- overwrite = FALSE loads data from the selected_dir directory parameter. 

```{r warning=FALSE, eval=FALSE}
# point to parent data directory
selected_dir <- "data/invitro/harmony_do_focus"
# Warning - this currently expects a temp_files data dir 
# Set overwrite to TRUE when underlying data dir has been changed
harmony_collection <- harmony_create_collection(dir = selected_dir, 
                                                overwrite = FALSE,
                                                temp_dir = "./temp_files",
                                                custom_rename = FALSE)
```

# Custom changes to data

## Update logic

- Revised data is then added to the collection as harmony_collection$all_plates_clean

```{r}
# potentially universal updates for class creator
harmony_collection$all_plates_clean <- harmony_collection$all_plates %>% 
  dplyr::rename(dose = concentration) %>%
  tibble::add_column(sex = NA, .after="individual") %>%
  dplyr::mutate(mouse = paste0(gsub('_','.',individual),"_", plate)) %>%
  dplyr::select(plate, row, column, dose, sex, individual, mouse, dplyr::everything())

# delete from dC20264_B1__20200928T12_28_55Measurement_1_e5 (Z)
# where rows are 6:8 and column is 9
# delete from dC20255_B2__20200926T13_00_37Measurement_2_e6 (Y)
# where rows are 1:6 and column is 10

# this needs a better standard / naming convention prior to analysis
excludes <- read.csv(text = "
dir,row,column
dC20264_B1__20200928T12_28_55Measurement_1_e5,6,9
dC20264_B1__20200928T12_28_55Measurement_1_e5,7,9
dC20264_B1__20200928T12_28_55Measurement_1_e5,8,9
dC20255_B2__20200926T13_00_37Measurement_2_e6,1,10
dC20255_B2__20200926T13_00_37Measurement_2_e6,2,10
dC20255_B2__20200926T13_00_37Measurement_2_e6,3,10
dC20255_B2__20200926T13_00_37Measurement_2_e6,4,10
dC20255_B2__20200926T13_00_37Measurement_2_e6,5,10
dC20255_B2__20200926T13_00_37Measurement_2_e6,6,10
")

harmony_collection$all_plates_clean <- dplyr::anti_join(harmony_collection$all_plates_clean, excludes)

# unique data cleaning
harmony_collection$all_plates_clean <- harmony_collection$all_plates_clean %>% 
  filter(
  dose != 0.5, 
  individual != '11_20',
  individual != '5346') %>%
  # incorrect pairs
  mutate(plate=ifelse((dir=='dC20222_B2_rerun__20200829T11_01_59Measurement_4_e9'), 'B2', plate)) %>%
  mutate(plate=ifelse((dir=='d20222_B1_Cal__20200825T12_23_59Measurement_3_e6'), 'B1', plate)) %>%
  mutate(plate=ifelse((dir=='d20223_B2_Cal__20200817T11_02_00Measurement_1_e9'), 'A2', plate)) %>%
  mutate(plate=ifelse((dir=='d20223_A1_Cal__20200817T17_03_14Measurement_2_e7'), 'A1', plate)) %>%
  mutate(plate=ifelse((dir=='d20223_A2_Cal__20200818T14_54_58Measurement_1_e6'), 'B2', plate)) %>%
  mutate(plate=ifelse((dir=='d20223_B1_Cal__20200827T10_48_20Measurement_1_e7'), 'B1', plate)) %>%
  # incomplete pairs / ensuring completion of plate column
  mutate(plate=ifelse((dir=='dC20236_B2__20200828T14_45_19Measurement_1_e8'), 'B2', plate)) %>%
  mutate(plate=ifelse((dir=='dC20255_B1__20200925T14_41_51Measurement_5_e6'), 'B1', plate)) %>%
  mutate(plate=ifelse((dir=='dC20255_B2__20200926T13_00_37Measurement_2_e6'), 'B2', plate)) %>%
  mutate(plate=ifelse((dir=='dC20236_B1__20200828T13_12_27Measurement_1_e7'), 'B1', plate)) %>%
  mutate(plate=ifelse((dir=='dC20255_B2_last3columns__20200928T09_33_41Measurement_1_e7'), 'B2', plate)) %>%
  mutate(plate=ifelse((dir=='dC20264_B1__20200928T12_28_55Measurement_1_e5'), 'B1', plate)) %>%
  mutate(plate=ifelse((dir=='dW20267_A1__20201002T10_37_07Measurement_1_e5'), 'A1', plate)) %>%
  mutate(plate=ifelse((dir=='dW20267_A2__20201004T15_35_00Measurement_1_e5'), 'A2', plate)) %>%
  mutate(plate=ifelse((dir=='dW20267_B1__20201004T12_18_54Measurement_2_e5'), 'B1', plate)) %>%
  mutate(plate=ifelse((dir=='dW20267_B2__20201002T12_32_44Measurement_2_e5'), 'B2', plate)) %>%
  mutate(dir=ifelse((dir=='dC20255_B2_last3columns__20200928T09_33_41Measurement_1_e7'), 'dC20255_B2__20200926T13_00_37Measurement_2_e6', dir)) %>%
  mutate(dir=ifelse((dir=='dC20264_B1_last3__20200928T13_26_45Measurement_1_e5'), 'dC20264_B1__20200928T12_28_55Measurement_1_e5', dir)) %>%
  # hardcode mapped replicates to paired IDs
  # this throws an error
  mutate(group_pair = LETTERS702[group_indices(., substr(plate,1,1), sub("\\_.*", "", dir))]) %>%
  mutate(
    fdose = factor(dose),
    fdir = factor(dir)
    )
```

## Notes on changes

- group_pairs with missing / removed well data
  - B (Individual 1882 instrument malfunction at corner of plate)
  - O (Individual 11_20 removed)
  - W (Individual 5346 removed due to naming error; data split between runs)
  - Y (Individual 5346 removed due to naming error; Individual 2213 instrument malfunction at corner of plate; Data split between runs)
  - Z (Data split between runs)

## Additional new columns

- this is needed for some subsequent plotting

```{r}
# integer bin_id logic
# binary integer ID per plate pair (zero or 1)
# this is slow / inefficient
harmony_collection$all_plates_clean <- iterate_bin_id(harmony_collection$all_plates_clean)
```

## Quick tables / plots

### Counts per feature

```{r}
table(harmony_collection$all_plates_clean$group_pair, harmony_collection$all_plates_clean$bin_id)
```

### Paired plot on example feature

```{r, fig.width=10, fig.height=10}
boxplot_pairs(y ='in_focus_nucleus_hoechst_33342_ser_hole_1_px_mean_per_well',
              data = harmony_collection$all_plates_clean,
              filename = NA)
```

# Saves

```{r}
# save over .rdata
saveRDS(harmony_collection, './harmony_do_focus.RData')
```

```{r}
# save corresponding .csv
# currently analyzing the following features
selected_features <- c(
  "nuclei_number_of_objects", 
  "in_focus_cell_area_mm2_mean_per_well",
  "in_focus_intensity_cell_hoechst_33342_mean_mean_per_well", 
  "in_focus_nucleus_hoechst_33342_ser_dark_1_px_mean_per_well", 
  "in_focus_nucleus_hoechst_33342_ser_edge_1_px_mean_per_well", 
  "in_focus_nucleus_hoechst_33342_ser_hole_1_px_mean_per_well",
  "in_focus_cell_mito_tracker_deep_red_ser_dark_1_px_mean_per_well", 
  "in_focus_cell_mito_tracker_deep_red_ser_edge_1_px_mean_per_well", 
  "in_focus_cell_mito_tracker_deep_red_ser_hole_1_px_mean_per_well",
  "in_focus_intensity_cell_mito_tracker_deep_red_mean_2_mean_per_well" 
  )
```

```{r}
summary(harmony_collection$all_plates_clean[,
     c("dir","row","column","dose","individual","mouse",
       selected_features)])
write.csv(harmony_collection$all_plates_clean[,
     c("dir","row","column","dose","individual","mouse",
       selected_features)], './harmony_do_focus_reduced.csv', row.names = FALSE) 
```