---
title: "Stage 2 DO Transformation"
author:
- name: Robert Sellers
  affiliation: The Jackson Laboratory
date: "`r paste('Last knit on:',format(Sys.time(), '%d %B %Y'))`"
abstract: |
latex_engine: pdflatex
output:
  html_document:
    fig_caption: yes
    force_captions: yes
    number_sections: true
    theme: lumen
    toc: yes
    toc_float: yes
    df_print: paged
    # code_folding: "hide"
  pdf_document:
    fig_caption: yes
    force_captions: yes
    highlight: pyments
    keep_tex: yes
    latex_engine: xelatex
    number_sections: no
    toc: no
    
---

Document/Process explained:

* The following documents batch correction and a summary analysis of features selected for analysis.
  
# Environment

- Point your working directory to the project root folder

```{r message=FALSE}
# env
setwd('~/Desktop/projects/jax_arsenic_toxicology_analysis/')
options(scipen = 999) # suppress sci notation
# warehouse for functions
source("scripts/custom_tools.R")
suppressPackageStartupMessages(require(tidyverse))
suppressPackageStartupMessages(require(kableExtra))
```

# Data

## Direct load from .rdata or .csv

- Use this if not working with Harmony software data or development.

```{r}
#harmony_collection <- readRDS('./harmony_do_focus.RData')
analytics_csv <- read.csv('./harmony_do_focus_reduced.csv')
```

# Data transformation / reduction

## Plate effects correction

```{r message = FALSE}
# formula for lme4
f_ <- formula(pheno ~ (1|individual) + (1|fdose) + (1|dir))

selected_features <- c(
  "nuclei_number_of_objects", 
  "in_focus_cell_area_mm2_mean_per_well",
  "in_focus_intensity_cell_hoechst_33342_mean_mean_per_well", 
  "in_focus_nucleus_hoechst_33342_ser_dark_1_px_mean_per_well", 
  "in_focus_nucleus_hoechst_33342_ser_edge_1_px_mean_per_well", 
  "in_focus_nucleus_hoechst_33342_ser_hole_1_px_mean_per_well",
  "in_focus_cell_mito_tracker_deep_red_ser_dark_1_px_mean_per_well", 
  "in_focus_cell_mito_tracker_deep_red_ser_edge_1_px_mean_per_well", 
  "in_focus_cell_mito_tracker_deep_red_ser_hole_1_px_mean_per_well",
  "in_focus_intensity_cell_mito_tracker_deep_red_mean_2_mean_per_well" 
  )

# create / backup df
features_corrected <- analytics_csv
features_corrected$fdose <- factor(features_corrected$dose)

# loop through the above selected features and apply plate(dir) batch correction
# batch correct script located in custom_scripts.R
for (feature in selected_features){
  features_corrected[feature] <- batch_correct(formula_ = f_, feature, features_corrected, r.eff = 'dir', target_col = 'dir', debug=FALSE)
  # assert zeros from negative adjustments
  features_corrected[feature] <- sapply(features_corrected[feature],function(x) ifelse(x<0, 0, x))
}
```

# Quick plots

## Raw with batch correct histograms

```{r warning = FALSE}
# wide to long format
raw_features_gathered <- gather(
  analytics_csv[selected_features],
  key = "feature", "value" = "value")
raw_features_gathered$sequence = 'raw'

corrected_features_gathered <- gather(
  features_corrected[selected_features],
  key = "feature", "value" = "value")
corrected_features_gathered$sequence = 'batch_corrected'

# cleaned names for plot
renamed <- setNames(c("num_nuclei","cell_area","hoecsht_intensity","hoechst_dark","hoechst_edge","hoechst_hole","mito_dark","mito_edge","mito_hole","mito_intensity"),selected_features)

ggplot(rbind(raw_features_gathered,corrected_features_gathered), aes(value, fill = sequence)) + 
    geom_histogram(
      bins = 200, 
      alpha = 0.5,
      position = "identity",
      #
      #binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3))
      ) + 
   theme(
     strip.background = element_rect(fill="lightblue", colour="black"),
     axis.text.y = element_text(size = 5),
     axis.text.x = element_text(size = 5,angle = 45, hjust = 1)
     ) +
    facet_grid(rows = vars(feature),scales = "free", labeller = labeller(feature = renamed)) + 
  scale_x_log10() +
  xlab("Response value density (axis log10)") +
  ylab("Count") +
  theme(strip.text.y = element_text(face = 'bold', angle = 0, size = 6)) 
```

## Summaries

```{r}
summary(features_corrected[selected_features])
```

## Missing data per feature by individual

```{r}
missing_by_feature <- features_corrected %>% 
  group_by(individual) %>% 
  summarise_all(~sum(is.na(.))) 
missing_mat <- missing_by_feature[c('individual',selected_features)][rowSums(missing_by_feature[selected_features]) > 0,]
kable(missing_mat, 
      linesep = "_", align = 'c',
      booktabs = T,
col.names = c("individual",c("num_nuclei","cell_area","hoecsht_intensity","hoechst_dark","hoechst_edge","hoechst_hole","mito_dark","mito_edge","mito_hole","mito_intensity"))) %>%
column_spec(1:11, width = "5cm")
```
# Saves

```{r}
# features_corrected.csv
write.csv(features_corrected, './harmony_do_focus_reduced_corrected.csv')
```