---
title: "In vivo growth data"
author:
- name: Robert Sellers
  affiliation: The Jackson Laboratory
date: "`r paste('Last knit on:',format(Sys.time(), '%d %B %Y'))`"
abstract: |
latex_engine: pdflatex
output:
  html_document:
    fig_caption: yes
    force_captions: yes
    number_sections: true
    theme: lumen
    toc: yes
    toc_float: yes
    df_print: paged
    # code_folding: "hide"
  pdf_document:
    fig_caption: yes
    force_captions: yes
    highlight: pyments
    keep_tex: yes
    latex_engine: xelatex
    number_sections: no
    toc: no
    
---

# Environment

```{r}
options(scipen = 999) # suppress sci notation

setwd('~/Desktop/projects/jax_arsenic_toxicology_analysis')

suppressPackageStartupMessages(require(tidyverse))
suppressPackageStartupMessages(require(readxl))
require(ggpubr)
require(zoo)

source('scripts/custom_tools.R')
```

# Load Data

- Does removed housing split data matter?

```{r}
housing <- read_excel(
    "./data/invivo/arsenic_mouse_bodyweight.xlsx",
    sheet = "housing",
    .name_repair = janitor::make_clean_names
  ) %>%
  separate(name, c("housing_var_a","id","housing_var_c"), sep='-') %>%
  select(-one_of("housing_var_a","housing_var_c")) 

bodyweights <- read_excel(
    "./data/invivo/arsenic_mouse_bodyweight.xlsx",
    sheet = "bodyweight",
    .name_repair = janitor::make_clean_names
  ) %>%
  dplyr::rename(generation = g) %>%
  mutate(
    id = factor(id),
    date_of_birth = as.Date(dob, tz = "EST")
  ) %>%
  left_join(housing) %>%
  group_by(housing_id) %>%
  arrange(id) %>%
  # adds a unique color column per id found
  group_modify(~ {
    .x %>%
      mutate(color_gen = sample(scales::hue_pal()(nrow(.))))
  }) %>%
  gather(week, bodyweight, bw8:bw52) %>%
  mutate(
   # extract numeric week integer
   weeks_integer = readr::parse_number(week),
   # adds week integer to real date
   week_as_date = dob + lubridate::weeks(weeks_integer),
   # round bodyweight to one decimal, NA warning is fine
   bodyweight = suppressWarnings(round(as.numeric(bodyweight),2))) %>%
  # handle duplicates
  group_by(id, weeks_integer) %>% 
  mutate(num_dups = n(), 
         dup_id = row_number()) %>% 
  ungroup() %>% 
  mutate(id = case_when(num_dups > 1 ~ paste0("D",dup_id,"_",id),
         TRUE ~ id)) %>%
  select(-one_of("week","dup_id","num_dups")) %>%
  mutate(fdose = factor(dose)) %>% 
  # add date ranking for normal date analysis
  group_by(wave, id) %>%
  mutate(date_order = order(order(week_as_date))) %>%
  #############
  # manual fixes
 # incorrect names or doses
  mutate(id=ifelse((id=='D2_10556'), '10556', id),
         id=ifelse((id=='D1_10658'),'10656', id),
         id=ifelse((id=='D2_10658'),'10658', id)) %>%
  mutate_cond(housing_id == 'H-89-19', dose = 1) %>%
  mutate_cond(housing_id == 'H-89-19', fdose = factor(1)) %>%
  mutate_cond(housing_id == 'H-85-19', dose = 0.1) %>%
  mutate_cond(housing_id == 'H-85-19', fdose = factor(0.1)) %>%
  # measurement errors
  # these data will get picked up for mean interpolation
  # do not attempt on first or last values in series
  mutate_cond(id =='10464' & date_order == 11, bodyweight = NA) %>%
  mutate_cond(id =='10261' & date_order == 8, bodyweight = NA) %>%
  mutate_cond(id =='10664' & date_order == 8, bodyweight = NA) %>%
  mutate_cond(id =='10189' & date_order == 10, bodyweight = NA) %>%
  mutate_cond(id =='10190' & date_order == 10, bodyweight = NA) %>%
  mutate_cond(id =='10309' & date_order == 8, bodyweight = NA) 
```

# Initiate fits on bodyweight data

- using what's available
- check https://www.statforbiology.com/nonlinearregression/usefulequations

```{r}
try_loess_lm <- function (df) {
  return(
    tryCatch({
      if (nrow(na.omit(df)) == 2 | nrow(na.omit(df)) == 1){
        lm(bodyweight ~ weeks_integer, data = na.omit(df))
      }else{
        loess(bodyweight ~ weeks_integer, data = na.omit(df)        ,control = loess.control(surface="direct"))
      }
      }
      , error=function(e) {
        paste0("failed with ", nrow(na.omit(df)))
      }
      )
    )
}

try_loess <- function (df) {
  return(
    tryCatch({
        loess(bodyweight ~ weeks_integer, data = na.omit(df)        ,control = loess.control(surface="direct"))
      }
      , error=function(e) {
        paste0("failed with ", nrow(na.omit(df)))
      }
      )
    )
}

try_predict <- function (x, df){
    return(tryCatch(suppressWarnings(predict(x, df)), error=function(e) NULL))
}

# generate loess models
# check list for NULL
loess_lm_models <- plyr::dlply(bodyweights, "id", function(df){
  suppressWarnings(try_loess_lm(df))
})

```

## Prepare causes_of_death data

```{r}
# provided via email 
deaths_supplement <- read_excel(
    "./data/invivo/arsenic_mouse_bodyweight.xlsx",
    sheet = "deaths_supplement",
    .name_repair = janitor::make_clean_names
  )  %>%
  mutate(
    death_exit_date = as.Date(death_exit_date, tz = "EST"),
    id = as.character(id)
  ) %>%
  dplyr::rename(week_as_date = death_exit_date) %>%
  filter(id %notin% c('100058','100059','10194',"100033","10391","10513","10515",  "10652","10686","10219","10316","10339")) %>%
  # need to fill dob 
  left_join(dplyr::distinct(bodyweights[,c("id","date_of_birth")],id,.keep_all = T)) %>%
  mutate(weeks_integer =as.numeric((week_as_date - date_of_birth)/7)) %>%
  dplyr::select(-date_of_birth)

deaths <- read_excel( 
    "./data/invivo/arsenic_mouse_bodyweight.xlsx",
    sheet = "deaths",
    .name_repair = janitor::make_clean_names
  )  %>%
  mutate(
    death_exit_date = as.Date(death_exit_date, tz = "EST"),
    cause_of_death = status
  ) %>%
  dplyr::rename(week_as_date = death_exit_date,
                weeks_integer = age_weeks) %>%
  separate(name, c("housing_var_a","id","housing_var_c"), sep='-') %>%
  select(-one_of("housing_var_a","housing_var_c", "status")) %>%
  left_join(housing)  %>%
  select(-one_of("sex","birth_date")) %>%
  dplyr::bind_rows(deaths_supplement) %>%
  mutate(
    est_method = 'documented'
    ) %>%
  plyr::join(dplyr::distinct(bodyweights[,c('dose','color_gen','id','fdose','sex','wave','generation')],id,.keep_all = T), by = 'id')


# creates dataset from final non-NA value encountered or documented deaths
cause_of_death <- bodyweights %>%
  mutate(
    est_method = "end_of_study"
    ) %>%
  drop_na(bodyweight) %>%
  bind_rows(deaths) %>%
  group_by(id) %>%
  filter(week_as_date == max(week_as_date)) %>%
  select(-date_order,date_of_birth,dob)

```

# Predictive analysis

- not running on 10656 or 10066
- loess models fitting on 67/

```{r}
# predict data where is documented death recorded
# WARNING: this is not finding duplicated models, but this wasn't necessary here for these 3/6 dups
# many model are returning as NULL
sub_models <- loess_lm_models[unlist(names(loess_lm_models) %in% cause_of_death[cause_of_death$est_method == 'documented', ]$id)]

# eval_fits_df <- plyr::ldply(sub_models, is.null) %>%
#   filter(V1 != FALSE)

# extracts predictive inputs from loess model list
sub_dates <- subset(cause_of_death, id %in% names(sub_models))[,c('id','weeks_integer')] %>%
  arrange(id)

# run predictions
predictions <- list()
for (i in seq_along(sub_models)) {
  pred_name <- names(sub_models)[i]
  # implement better error catching here
  predictions[pred_name] <- try_predict(
    x = sub_models[[i]], 
    df = data.frame("weeks_integer"= sub_dates$weeks_integer[i])
  )
}

# cleanup
pred_df <- stack(predictions) %>%
   dplyr::rename(id = ind, predicted_deaths = values)

# apply predictions to dataframe
cause_of_death <- cause_of_death %>%
  plyr::join(pred_df, by = 'id') %>%
  mutate_cond(is.na(bodyweight), bodyweight = predicted_deaths) %>%
  mutate(week_as_date = as.Date(week_as_date, tz = "EST"),
         bodyweight_fit = bodyweight)
```

# Labels for ggplot

```{r}
# list no data
na_list <- bodyweights %>%
  group_by(id,housing_id,wave,generation,dose,fdose, sex) %>% 
  summarise_each(funs(100*mean(is.na(.)))) %>%
  filter(bodyweight == 100) %>%
  group_by(housing_id,sex,wave,dose,fdose,generation) %>%
  summarise(label_nodata = paste("no data:", id, collapse = "\n"),
            label_group = paste(paste0("wave:",wave),housing_id, paste0("wave:",sex), paste0("wave:",dose), collapse = "\n"))
```

# Simple impute / interpolation

- operates on NA between existing values

```{r}
bodyweights <-  bodyweights %>%
  group_by(id) %>%
  mutate(bodyweight_test = na.approx(bodyweight, na.rm=FALSE)) %>%
  mutate_cond(is.na(bodyweight) & !is.na(bodyweight_test), bodyweight = bodyweight_test) %>%
  dplyr::select(-bodyweight_test)
```

# Extract loess model fits

- this is where data may disappear
- need to track null / single value bodyweights prior to this

```{r}
bw_original <- plyr::ldply(loess_lm_models, function(mod){
  # first construct a data.frame for entire list of models x / y
  if (class(mod) == "loess"){
    data.frame("weeks_integer" = mod$x, "bodyweight" = mod$fitted)
  }
  }) %>%
  mutate(tmp_id = gsub(".*_", "", id)) %>%
  left_join(housing, by = c("tmp_id" = "id")) %>%
  right_join(bodyweights, 
             by=c("id","weeks_integer"),
             suffix = c("_fit", "_recorded")) %>%
  select(-one_of("tmp_id","housing_id_fit","dob")) %>% 
  # needed?
  mutate(
    week_as_date = as.Date(week_as_date, tz = "EST")
  ) %>%
  dplyr::rename(housing_id = housing_id_recorded,
                bodyweight = bodyweight_recorded) %>%
  group_by(housing_id) %>%
  # add baseline values where fits incomplete
  mutate_cond(is.na(bodyweight_fit), bodyweight_fit = bodyweight)
  ######optional######
   # sample_n_groups(10) 
  ####################

# add deaths to fit prior to smooth line
bw_fit_w_estimates <- rbind(bw_original,filter(cause_of_death)) %>%
  group_by(id) %>%
  mutate(min_date = min(week_as_date)) %>%
  group_by(id) %>%
  mutate(weeks_integer = (week_as_date - min_date)/7) %>%
  group_by(id) %>%
  tidyr::fill(date_of_birth) %>%
  tidyr::fill(date_of_birth, .direction = "up")
```

## Group fit from predicted
- very hacky code 

```{r}
# k <- bw_original %>% group_by(wave) %>% 
# summarize(date = sort(unique(week_as_date),collapse=", ")) %>%
#   group_split(wave)
# 
# k2 <- do.call(cbind, k) 
# valid_column_names <- make.names(names=names(k2), unique=TRUE, allow_ = TRUE)
# names(k2) <- valid_column_names
# 
# k3 <- k2 %>% 
#   dplyr::select(starts_with("date")) %>%
#   rename_all(funs(c("1", "2","3","4","5","6","7","8"))) %>% 
#   mutate(
#     date_order=as.integer(row_number())
#     ) %>%
#   gather(wave,week_as_date, 1:8)
# 
# k4 <- bw_fit_w_estimates %>% 
#   expand(date_order = seq(1:12)) %>%
#   left_join(k3, by='date_order')
# 
# k5 <- k4 %>%
#   group_by(housing_id,date_order,wave,week_as_date) %>%
#   expand(dose = unique(bw_fit_w_estimates$dose),
#          sex = c("M","F")) 
# 
# bw_fit_w_estimates_means <- bw_fit_w_estimates %>%
#   drop_na(bodyweight) %>%
#   group_by(wave, sex, dose, week_as_date) %>%
#   summarize(bodyweight = mean(bodyweight))
# 
# k6 <- bw_fit_w_estimates_means %>%
#   mutate(wave = as.character(wave)) %>%
#   semi_join(k5, by = c("wave","sex", "dose")) %>%
#   mutate(id = paste0(sex,"-",dose,"-",wave),
#          wave = as.numeric(wave))
# 
# k7 <- left_join(k6, bw_fit_w_estimates, by = c('wave', 'sex','dose','week_as_date')) %>%
#   select(wave,sex,dose,week_as_date,bodyweight.x,id.x,housing_id) %>%
#   dplyr::rename(bodyweight = bodyweight.x, id = id.x) %>%
#   distinct()
# 
# rm(k,k2,k3,k4,k5,k6)
```

## reverse engineer K7

 - more hacky code

```{r}
# lookup_agg <- bw_fit_w_estimates %>%
#   drop_na(bodyweight) %>%
#   group_by(sex,dose,date_order) %>%
#   summarize(bodyweight = mean(bodyweight))
# 
# # This works but the CIs are junk
# k8 <- k7 %>%
#   select(-one_of("bodyweight")) %>%
#   group_by(wave, id, housing_id) %>%
#   mutate(date_order = order(order(week_as_date))) %>%
#   left_join(lookup_agg, by = c('sex','dose','date_order')) %>% 
#   mutate(color_gen = ifelse(sex == 'M', 'blue', 'pink'))
```

## Fits plot

 - not knit - look for PDF

```{r eval = FALSE}

pdf("./output/bodyweight/mice_by_cage_v26.pdf")
# use n_pages on ggplot object on one page to estimate loop size
#ggforce::n_pages()

for (i in 1:ceiling(length(unique(bw_fit_w_estimates$housing_id))/5)){
  print(paste0("producing  plot page #", i))
  print(
  ggplot(arrange(bw_fit_w_estimates, sex, wave, housing_id, dose), 
    aes(x = week_as_date, y = bodyweight, group = id)) +
  geom_path(aes(color = color_gen), 
            size = 0.5,
            linetype = 'dotdash') +
      geom_text(data = na_list, x=-Inf, y=Inf, aes(group = housing_id,label = label_nodata),hjust = 0, vjust = 1, size=3) +
      # geom_text(data = na_list, x=Inf, y=-Inf, aes(group = housing_id,label = label_group),hjust = 0, vjust = 1, size=3) +
  scale_shape_manual(values = c(16, 1)) +   # doc, end
    # smooths on lm results
  geom_smooth(data = filter(bw_fit_w_estimates,id %in% names(which(sapply(loess_lm_models, function(x) (class(x) == "lm")),TRUE))),
              method = 'lm',
              aes(x = week_as_date,
                  y = bodyweight,
                  color = color_gen),
              linetype = 'solid',
              se = FALSE,
              size=0.5) +
    # smooths on loess results
  geom_smooth(data = filter(bw_fit_w_estimates,id %in% names(which(sapply(loess_lm_models, function(x) (class(x) == "loess")),TRUE))),
              method = 'loess',
              aes(x = week_as_date,
                  y = bodyweight,
                  color = color_gen),
              linetype = 'solid',
              se = FALSE,
              size=0.5) +
  geom_point(data = bw_fit_w_estimates, aes(x = week_as_date, y = bodyweight), size=0.15) +
  geom_point(data = cause_of_death,
             size = 2, show.legend = FALSE,
             aes(x = as.Date(week_as_date), #weird
                 y = bodyweight,
                 shape = est_method,
                 color = color_gen)
             ) +
  ggrepel::geom_text_repel(
    aes_q(label = ~paste(id,"\n", 
          ifelse(is.na(cause_of_death),"",paste0(sprintf("%s",cause_of_death),"\n")),
          ifelse(is.na(comments),"",sprintf("%s",comments))), 
           color = as.character(cause_of_death$color_gen)),
           fontface = "bold",
           direction = 'both',
           nudge_x = 25,
           bg.color = "white",
           bg.r = 0.2,
           force_pull = 3,
           segment.colour = 'black',
           min.segment.length = unit(0, 'lines'),
           data = cause_of_death,
           size = 2)  + 
  ######### setting x limits, nudging to the right #########
ggforce::facet_wrap_paginate(~ housing_id + sex + dose + wave, 
                               ncol = 1, 
                               nrow = 5,
                               page  = i,
                               strip.position="right",
                               scales = "free_y",
                               labeller = labeller( wave = label_both, housing_id = label_both, dose = label_both, sex = label_both)) + 
  scale_x_date(date_labels = "%b-%y", date_breaks = '1 month',minor_breaks=NULL) +
  ylab("Bodyweights (grams)") +
  xlab("Arsenic Dose (ppm)") +
  theme(
    strip.text = element_text(size=8),
    panel.border = element_rect(fill=NA, colour="grey40"),
    axis.text.x = element_text(size = 6, angle = 45, vjust = 1.1, hjust=1),
    axis.title.x=element_blank()
  ) + guides(colour = FALSE, shape = guide_legend(),
  linetype = guide_legend())
  )
}
dev.off()
```

## QA from visual analysis

- manual evaluation of irregular data

```{r}
bodyweight_qa <- read_excel(
    "./data/invivo/bodyweights_qa.xlsx",
    sheet = "Sheet1",
    .name_repair = janitor::make_clean_names
  )
# test same data
check_missing_deaths <- subset(deaths, id %in% unique(filter(bodyweight_qa, problem == 'dne')$id))

check_oneval <- subset(bodyweights, id %in% unique(filter(bodyweight_qa, problem == '1_val')$id)) %>%
  drop_na(bodyweight)
check_oneval_deaths <- subset(deaths, id %in% unique(filter(bodyweight_qa, problem == '1_val')$id)) %>%
  drop_na(bodyweight)

check_missing_data <- subset(bodyweights, id %in% unique(filter(bodyweight_qa, problem == 'missing_data')$id)) %>%
  drop_na(bodyweight)
check_missing_data_deaths <- subset(deaths, id %in% unique(filter(bodyweight_qa, problem == 'missing_data')$id)) %>%
  drop_na(bodyweight)
```

## ttest + anova for  bodyweight decrease

- check out https://rpkgs.datanovia.com/rstatix/

```{r}
bw_fit_w_estimates_quickfix <- bw_fit_w_estimates %>%
  drop_na(fdose)

bw_fit_w_estimates_quickfix %>%
  group_by(sex,fdose) %>%
  get_summary_stats(bodyweight, type = "mean_sd") 


# see if significance in reduction of bodyweight @ 10 ppm
# 10 < 1

dose_pairs <- expand.grid(dose_1 = unique(bw_fit_w_estimates_quickfix$dose), dose_2 = unique(bw_fit_w_estimates_quickfix$dose)) %>%
  mutate(p = NA,
         stderr = NA) %>%
  filter(dose_1 != dose_2)

for (i in 1:nrow(dose_pairs)){

    g_a <- subset(bw_fit_w_estimates, dose==dose_pairs[i,]$dose_1)
  g_b <- subset(bw_fit_w_estimates, dose==dose_pairs[i,]$dose_2)
ttest_temp <- t.test(
      g_a$bodyweight,
      g_b$bodyweight, 
      alternative = "less"
     )
    dose_pairs[i,"p"] <- ttest_temp$p.value
    dose_pairs[i,"stderr"] <- ttest_temp$stderr
}

dose_pairs <- dose_pairs %>%
  mutate(p = round(p,5),
         stderr = round(stderr,4)) %>% 
  arrange(p)

# Anova
anova_sex <- aov(bodyweight ~ fdose + sex + sex:fdose, data = bw_fit_w_estimates)
# summary(anova_sex)
coefficients(anova_sex)
```

## Bodyweight avg plot

```{r}
small_axis_theme = theme(
  axis.title.x = element_text(size = 8),
  axis.text.x = element_text(size = 6),
  axis.text.y = element_text(size = 6),
  axis.title.y = element_text(size = 8))

cc_months <- scales::seq_gradient_pal("blue", "red", "Lab")(seq(0,1,length.out=12))

cc_fdose <- scales::seq_gradient_pal("green", "red", "Lab")(seq(0,1,length.out=5))

cc_initweight <- scales::seq_gradient_pal("green", "red", "Lab")(seq(0,1,length.out=10))

bw_original <- bw_original %>%
  group_by(id) %>%
  mutate(initial_bw    = first(bodyweight), 
         f_dateorder = factor(date_order),
         initial_bw_num =as.numeric(initial_bw))

# table(bw_original$initial_bw, bw_original$dose)
bw_original$initial_bw <- as.factor(Hmisc::cut2(bw_original$initial_bw, g=10))

############################
df_all <- CreateAllFacet(bw_original, "sex")
a <- ggplot()+
  geom_smooth(data = filter(df_all, !is.na(bodyweight)), 
              aes(x = f_dateorder,
                  y = bodyweight,
                  group=fdose,
                  color = fdose
                  ),
              linetype = 'solid',
              size = 0.5,
              span = 0.6,
              se = FALSE
              ) + facet_wrap(~ facet,dir = 'h') + ggtitle("sex x dose") +     
    scale_colour_manual(values=cc_fdose) + small_axis_theme 
a <- addSmallLegend_8(a)
############################
b <- ggpubr::ggtexttable(subset(dose_pairs, p < 0.3), theme = ggpubr::ttheme("classic")) %>%
 tab_add_title(text = 't-tests; dose pairs < 0.05 asc', padding = unit(1.5, "line")) 

b <- table_cell_bg(b, row = 3:8, column = 2:5,
                      fill="darkolivegreen1")

############################
c <- ggplot()+
  geom_smooth(data =filter(bw_original, !is.na(bodyweight)),
              aes(x = fdose,
                  y = bodyweight,
                  group = interaction(sex,f_dateorder),
                  color = f_dateorder
              ),
              size = 0.5,
              span = 0.6,
              se = FALSE
              ) + facet_wrap(~ sex,dir = 'h') +ggtitle("sex:dateorder interaction")  +
    scale_colour_manual(values=cc_months) + small_axis_theme
c <- addSmallLegend_8(c)
############################
d <- ggplot()+
  geom_smooth(data = filter(bw_original, !is.na(bodyweight) & fdose %in% c(0,10)), 
              aes(x = f_dateorder,
                  y = bodyweight,
                  group = interaction(initial_bw, fdose),
                  color = initial_bw
                  ),
              linetype = 'solid',
              size = 0.5,
              span = 0.6,
              se = FALSE
              ) + ggtitle("sex x dose:initial_bw interaction") +     
    scale_colour_manual(values=cc_initweight) + facet_grid(~ fdose~sex) + small_axis_theme

d <- addSmallLegend_8(d)

pdf("./output/bodyweight/eval_bw_4plot_v1.pdf")
ggpubr::ggarrange(a, b, c, d, common.legend = F, ncol = 2, nrow = 2) 
dev.off()
```


```{r}
 # ggplot()+
 #  geom_smooth(data = filter(bw_original, !is.na(bodyweight)), 
 #              aes(x = factor(date_order),
 #                  y = bodyweight,
 #                  group= interaction(sex,dose),
 #                  color = sex
 #                  ),
 #              linetype = 'solid',
 #              size = 0.4,
 #             # span = 0.6,
 #              alpha = 0.6,
 #              se = FALSE
 #              ) + facet_wrap(~ initial_bw ,dir = 'h', scales = "fixed")+ ggtitle("sex:dose interaction with initial weight bins")
ggplot()+
  geom_smooth(data =filter(bw_original, !is.na(bodyweight)), 
              
              aes(x = fdose,
                  y = bodyweight,
                  group = interaction(sex,f_dateorder),
                  color = date_order,
                  linetype = sex
                  ),
             # linetype = 'solid',
              size = 0.5,
              span = 0.6,
              se = FALSE
              ) + ggtitle("sex:dateorder interaction") + 
  viridis::scale_color_viridis( option = "D", "Month+1")

```

# Variance analysis

## Initial boxplot

```{r}
#https://rpkgs.datanovia.com/ggpubr/reference/ggpar.html
p <- ggpubr::ggboxplot(filter(bw_fit_w_estimates, !is.na(dose)), x = "dose", y = "bodyweight", color='sex')
#ggpubr::ggpar(p, legend = "none")
ggpubr::ggpar(p)
```

## Dose factor

```{r}
anova_fdose <- aov(bodyweight ~ fdose, data = bw_fit_w_estimates)
summary(anova_fdose)
plot(anova_fdose,1)
plot(anova_fdose,2)
broom::tidy(TukeyHSD(anova_fdose, "fdose", conf.level = 0.95))
```

## Dose factor with id

```{r}
anova_id <- aov(bodyweight ~ fdose + sex + sex:fdose + sex:fdose:factor(date_order) , data = bw_fit_w_estimates)
anova_id
anova_id$coefficients

# coefs <- data.frame("id" = gsub("[^0-9.-]", "",names(anova_id$coefficients)), "variance" =gsub("[^0-9.-]", "",anova_id$coefficients)
# ) %>%
#   mutate(variance = as.numeric(variance))%>%
#         inner_join(., bw_fit_w_estimates, by = "id") %>%
#   distinct(id, wave, dose, variance,sex)
# broom::tidy(TukeyHSD(anova_id, "id", conf.level = 0.95))
# plot(anova_id,1)
# plot(anova_id,2)
DT::datatable(broom::tidy(TukeyHSD(anova_id, "date_order", conf.level = 0.95)))
```

## Dose factor with sex and sex:fdose interaction

```{r}
anova_sex <- aov(bodyweight ~ sex + sex:fdose , data = bw_fit_w_estimates)



# is significant on at least one dose
summary(anova_sex)

DT::datatable(broom::tidy(TukeyHSD(anova_sex, "sex:fdose", conf.level = 0.95)))
plot(anova_sex,1)
plot(anova_sex,2)
```

## Dose factor with initial bw

```{r}
# anova_sex <- aov(bodyweight ~ sex + sex:fdose , data = bw_fit_w_estimates)
```