---
title: "In vivo growth data"
author:
- name: Robert Sellers
  affiliation: The Jackson Laboratory
date: "`r paste('Last knit on:',format(Sys.time(), '%d %B %Y'))`"
abstract: |
latex_engine: pdflatex
output:
  html_document:
    fig_caption: yes
    force_captions: yes
    number_sections: true
    theme: lumen
    toc: yes
    toc_float: yes
    df_print: paged
    # code_folding: "hide"
  pdf_document:
    fig_caption: yes
    force_captions: yes
    highlight: pyments
    keep_tex: yes
    latex_engine: xelatex
    number_sections: no
    toc: no
    
---

# Environment

```{r}
options(scipen = 999) # suppress sci notation

setwd('~/Desktop/projects/jax_arsenic_toxicology_analysis')

suppressPackageStartupMessages(require(tidyverse))
suppressPackageStartupMessages(require(readxl))
require(ggpubr)

source('scripts/custom_tools.R')
```

# Load Data

- Does removed housing split data matter?

```{r}
housing <- read_excel(
    "./data/invivo/arsenic_mouse_bodyweight.xlsx",
    sheet = "housing",
    .name_repair = janitor::make_clean_names
  ) %>%
  separate(name, c("housing_var_a","id","housing_var_c"), sep='-') %>%
  select(-one_of("housing_var_a","housing_var_c")) 

bodyweights <- read_excel(
    "./data/invivo/arsenic_mouse_bodyweight.xlsx",
    sheet = "bodyweight",
    .name_repair = janitor::make_clean_names
  ) %>%
  dplyr::rename(generation = g) %>%
  mutate(
    id = factor(id),
    date_of_birth = as.Date(dob, tz = "EST")
  ) %>%
  left_join(housing) %>%
  group_by(housing_id) %>%
  # adds a unique color column per id found
  group_modify(~ {
    .x %>%
      mutate(color_gen = sample(scales::hue_pal()(nrow(.))))
  }) %>%
  gather(week, bodyweight, bw8:bw52) %>%
  mutate(
   # extract numeric week integer
   weeks_integer = readr::parse_number(week),
   # adds week integer to real date
   week_as_date = dob + lubridate::weeks(weeks_integer),
   # round bodyweight to one decimal, NA warning is fine
   bodyweight = suppressWarnings(round(as.numeric(bodyweight),1))) %>%
  # handle duplicates
  group_by(id, weeks_integer) %>% 
  mutate(num_dups = n(), 
         dup_id = row_number()) %>% 
  ungroup() %>% 
  mutate(id = case_when(num_dups > 1 ~ paste0("D",dup_id,"_",id),
         TRUE ~ id)) %>%
  select(-one_of("week","dup_id","num_dups")) %>%
  mutate(fdose = factor(dose)) %>% 
# add date ranking for normal date analysis
  group_by(wave, id) %>%
  mutate(date_order = order(order(week_as_date)))

deaths <- read_excel(
    "./data/invivo/arsenic_mouse_bodyweight.xlsx",
    sheet = "deaths",
    .name_repair = janitor::make_clean_names
  )  %>%
  mutate(
    death_exit_date = as.Date(death_exit_date, tz = "EST"),
    cause_of_death = status,
    bodyweight = NA
  ) %>%
  dplyr::rename(week_as_date = death_exit_date,
                weeks_integer = age_weeks) %>%
  separate(name, c("housing_var_a","id","housing_var_c"), sep='-') %>%
  select(-one_of("housing_var_a","housing_var_c", "status")) %>%
  left_join(housing)  %>%
  select(-one_of("sex","birth_date")) 
```

# Initiate fits on bodyweight data

-check https://www.statforbiology.com/nonlinearregression/usefulequations

```{r}
try_loess <- function (df) {
  return(
    tryCatch(
    loess(bodyweight ~ weeks_integer, data = df,control=loess.control(surface="direct"))
    , error=function(e) {
      NULL
    }
    ))
}

try_predict <- function (x, df){
    return(tryCatch(suppressWarnings(predict(x, df)), error=function(e) NULL))
}

# generate loess models
# check list for NULL
loess_models <- plyr::dlply(bodyweights, "id", function(df){
  suppressWarnings(try_loess(df))
})
```

## Prepare causes_of_death data

```{r}
# creates dataset from final non-NA value encountered
cause_of_death <-  bodyweights %>% 
  group_by(id) %>%
  filter(weeks_integer == max(weeks_integer)) %>%
  select(-one_of("weeks_integer")) %>%
  mutate(
    est_method = "end_of_study"
    ) %>%
  bind_rows(deaths)
```

# Predictive analysis

```{r}
# predict data where is documented death recorded
# WARNING: this is not finding duplicated models, but this wasn't necessary here for these 3/6 dups
# many model are returning as NULL
sub_models <- loess_models[unlist(names(loess_models) %in% cause_of_death[is.na(cause_of_death$est_method), ]$id)]

# extracts predictive inputs from loess model list
sub_dates <- subset(cause_of_death, id %in% names(sub_models))[,c('id','weeks_integer')] %>%
  na.omit() %>%
  arrange(id)

# run predictions
predictions <- list()
for (i in seq_along(sub_models)) {
  pred_name <- names(sub_models)[i]
  # implement better error catching here
  predictions[pred_name] <- try_predict(
    x = sub_models[[i]], 
    df = data.frame("weeks_integer"= sub_dates$weeks_integer[i])
  )
}
#unsuccessful models
#sapply(sub_models, function(y) sum(length(which(is.null(y)))))

# cleanup
pred_df <- stack(predictions) %>%
   dplyr::rename(id = ind, bodyweight = values)

```

## Rebuilding deaths from predicted data

```{r}
# apply predictions to dataframe
# zeros seem to be applying
# ONLY RUN ON NA / NOT END_OF_STUDY
deaths_rebuild <- cause_of_death %>%
  filter(is.na(est_method)) %>%
  mutate(est_method = 'documented')

deaths_rebuild$bodyweight[match(pred_df$id, deaths_rebuild$id)] <- pred_df$bodyweight

# rebuild all_deaths
all_deaths <- filter(cause_of_death, !is.na(est_method)) %>%
  bind_rows(deaths_rebuild) %>%
  group_by(id) %>%
  fill(wave, generation, color_gen, dose) 
```

# Extract loess model fits

```{r}
bw_original <- plyr::ldply(loess_models, function(mod){
  # first construct a data.frame for entire list of models x / y
  data.frame("weeks_integer" = mod$x, "bodyweight" = mod$fitted)
  }) %>%
  mutate(tmp_id = gsub(".*_", "", id)) %>%
  left_join(housing, by = c("tmp_id" = "id")) %>%
  right_join(bodyweights, 
             by=c("id","weeks_integer"),
             suffix = c("_fit", "_recorded")) %>%
  select(-one_of("tmp_id","housing_id_fit","dob")) %>% 
  # needed?
  mutate(
    week_as_date = as.Date(week_as_date, tz = "EST")
  ) %>%
  dplyr::rename(housing_id = housing_id_recorded,
                bodyweight = bodyweight_recorded) %>%
  group_by(housing_id) #%>% 
  ######optional######
   # sample_n_groups(10) 
  ####################

bw_fit <- bw_original %>%
  select(-one_of("bodyweight_fit")) 

deaths_cleaned <- all_deaths %>%
  filter(id %in% unique(bw_original$id)) %>%
  mutate(week_as_date = as.Date(week_as_date, tz = "EST")) %>%
  drop_na(bodyweight) %>%
  fill(sex,dob,fdose,dose) 
# %>%
#   group_by(id) %>%
#   arrange(est_method) %>%
#   slice(1)

# add deaths to fit prior to smooth line
bw_fit_w_estimates <- rbind(bw_fit,filter(deaths_cleaned,est_method == 'documented')) %>%
  fill(sex,dose,date_order)
```

## QA dataset

```{r}
# qa time
bw_hs <- as.data.frame(table(bodyweights$wave, bodyweights$housing_id, bodyweights$id)) %>%
  filter(Freq != 0) %>%
  dplyr::rename(
    wave = Var1,
    housing_id = Var2,
    id = Var3,
    count = Freq
  )
```

## Group fit from predicted
- very hacky code 

```{r}
k <- bw_original %>% group_by(wave) %>% 
summarize(date = sort(unique(week_as_date),collapse=", ")) %>%
  group_split(wave)

k2 <- do.call(cbind, k) 
valid_column_names <- make.names(names=names(k2), unique=TRUE, allow_ = TRUE)
names(k2) <- valid_column_names

k3 <- k2 %>% 
  dplyr::select(starts_with("date")) %>%
  rename_all(funs(c("1", "2","3","4","5","6","7","8"))) %>% 
  mutate(
    date_order=as.integer(row_number())
    ) %>%
  gather(wave,week_as_date, 1:8)

k4 <- bw_fit_w_estimates %>% 
  expand(date_order = seq(1:12)) %>%
  left_join(k3, by='date_order')

k5 <- k4 %>%
  group_by(housing_id,date_order,wave,week_as_date) %>%
  expand(dose = unique(bw_fit_w_estimates$dose),
         sex = c("M","F")) 

bw_fit_w_estimates_means <- bw_fit_w_estimates %>%
  drop_na(bodyweight) %>%
  group_by(wave, sex, dose, week_as_date) %>%
  summarize(bodyweight = mean(bodyweight))

k6 <- bw_fit_w_estimates_means %>%
  mutate(wave = as.character(wave)) %>%
  semi_join(k5, by = c("wave","sex", "dose")) %>%
  mutate(id = paste0(sex,"-",dose,"-",wave),
         wave = as.numeric(wave))

k7 <- left_join(k6, bw_fit_w_estimates, by = c('wave', 'sex','dose','week_as_date')) %>%
  select(wave,sex,dose,week_as_date,bodyweight.x,id.x,housing_id) %>%
  dplyr::rename(bodyweight = bodyweight.x, id = id.x) %>%
  distinct()

rm(k,k2,k3,k4,k5,k6)
```

## reverse engineer K7

 - more hacky code

```{r}
lookup_agg <- bw_fit_w_estimates %>%
  drop_na(bodyweight) %>%
  group_by(sex,dose,date_order) %>%
  summarize(bodyweight = mean(bodyweight))

# This works but the CIs are junk
k8 <- k7 %>%
  select(-one_of("bodyweight")) %>%
  group_by(wave, id, housing_id) %>%
  mutate(date_order = order(order(week_as_date))) %>%
  left_join(lookup_agg, by = c('sex','dose','date_order')) %>% 
  mutate(color_gen = ifelse(sex == 'M', 'blue', 'pink'))
```

## Fits plot

 - not knit - look for PDF

```{r eval = FALSE}

pdf("./output/bodyweight/mice_by_cage_v12.pdf")
# use n_pages on ggplot object on one page to estimate loop size
#ggforce::n_pages()

for (i in 1:ceiling(length(unique(bw_original$housing_id))/4)){
  print(
  ggplot(arrange(bw_original, sex, wave, housing_id, dose), 
    aes(x = week_as_date, y = bodyweight, group = id)) +
     # geom_vline(aes(xintercept=week_as_date),
     #            linetype=5, colour="black",  size = .1) +
    
    # xlim(min(bw_original$week_as_date), max(bw_original$week_as_date)+ months(3)) +
  ################################
  geom_smooth(data = k8, 
              aes(x = week_as_date,
                  y = bodyweight
                  ),
              linetype = 'solid',
              color = 'black',
              se = FALSE,
              span = 0.6
              ) +
  geom_path(aes(color = color_gen), 
            size = 0.5,
            linetype = 'dotdash') +
  geom_smooth(data = bw_fit_w_estimates,
              aes(x = week_as_date,
                  y = bodyweight,
                  color=color_gen),
              linetype = 'solid',
              se = FALSE,
              size=0.5) +
  geom_point(data = bw_original, aes(x = week_as_date, y = bodyweight), size=0.15) +
  geom_point(data = deaths_cleaned,
             size = 2, show.legend = FALSE,
             aes(x = as.Date(week_as_date), #weird
                 y = bodyweight,
                 shape = est_method,
                 color = color_gen)
             ) +
  scale_shape_manual(values = c(16, 1)) +   # doc, end
  ggrepel::geom_text_repel(
    aes_q(label = ~paste(id,"\n", 
          ifelse(is.na(cause_of_death),"",paste0(sprintf("%s",cause_of_death),"\n")),
          ifelse(is.na(comments),"",sprintf("%s",comments))), 
           color = as.character(deaths_cleaned$color_gen)),
           fontface = "bold",
           direction = 'both',
           nudge_x = 25,
           #arrow = grid::arrow(),
           force_pull = 3,
           segment.colour = 'black',
           #min.segment.length = unit(0, 'lines'),
           data = deaths_cleaned,
           size = 2)  + 
  ######### setting x limits, nudging to the right #########
ggforce::facet_wrap_paginate(~ sex + dose + wave + housing_id, 
                               ncol = 1, 
                               nrow = 4,
                               page  = i,
                               strip.position="right",
                               scales = "free_y",
                               labeller = labeller( wave = label_both, housing_id = label_both, dose = label_both, sex = label_both)) + 
  scale_x_date(date_labels = "%b-%y", date_breaks = '1 month',minor_breaks=NULL) +
  ylab("Bodyweights (grams)") +
  xlab("Arsenic Dose (ppm)")
  # ylim(min(bw_original$bodyweight_fit,bw_original$bodyweight),max(bw_original$bodyweight_fit,bw_original$bodyweight)) +
  #xlab("                                          < 2019 | 2020 >                                                        | 2021")+
  theme(
    strip.text = element_text(size=8),
    panel.border = element_rect(fill=NA, colour="grey40"),
    axis.text.x = element_text(size = 6, angle = 45, vjust = 1.1, hjust=1),
    axis.title.x=element_blank()
  ) + guides(colour = FALSE, shape = guide_legend(),
  linetype = guide_legend())

  )
}
dev.off()
```

## ttest + anova for  bodyweight decrease

- check out https://rpkgs.datanovia.com/rstatix/

```{r}
bw_fit_w_estimates_quickfix <- bw_fit_w_estimates %>%
  drop_na(fdose)

bw_fit_w_estimates_quickfix %>%
  group_by(sex,fdose) %>%
  get_summary_stats(bodyweight, type = "mean_sd") 


# see if significance in reduction of bodyweight @ 10 ppm
# 10 < 1

dose_pairs <- expand.grid(dose_1 = unique(bw_fit_w_estimates_quickfix$dose), dose_2 = unique(bw_fit_w_estimates_quickfix$dose)) %>%
  mutate(p = NA,
         stderr = NA) %>%
  filter(dose_1 != dose_2)

for (i in 1:nrow(dose_pairs)){

    g_a <- subset(bw_fit_w_estimates, dose==dose_pairs[i,]$dose_1)
  g_b <- subset(bw_fit_w_estimates, dose==dose_pairs[i,]$dose_2)
ttest_temp <- t.test(
      g_a$bodyweight,
      g_b$bodyweight, 
      alternative = "less"
     )
    dose_pairs[i,"p"] <- ttest_temp$p.value
    dose_pairs[i,"stderr"] <- ttest_temp$stderr
}

dose_pairs <- dose_pairs %>%
  mutate(p = round(p,5),
         stderr = round(stderr,4)) %>% 
  arrange(p)

# Anova
anova_sex <- aov(bodyweight ~ fdose + sex + sex:fdose, data = bw_fit_w_estimates)
# summary(anova_sex)
coefficients(anova_sex)
```

## Bodyweight avg plot

```{r}
small_axis_theme = theme(
  axis.title.x = element_text(size = 8),
  axis.text.x = element_text(size = 6),
  axis.text.y = element_text(size = 6),
  axis.title.y = element_text(size = 8))

cc_months <- scales::seq_gradient_pal("blue", "red", "Lab")(seq(0,1,length.out=12))

cc_fdose <- scales::seq_gradient_pal("green", "red", "Lab")(seq(0,1,length.out=5))

cc_initweight <- scales::seq_gradient_pal("green", "red", "Lab")(seq(0,1,length.out=10))

bw_original <- bw_original %>%
  group_by(id) %>%
  mutate(initial_bw    = first(bodyweight), 
         f_dateorder = factor(date_order),
         initial_bw_num =as.numeric(initial_bw))

# table(bw_original$initial_bw, bw_original$dose)
bw_original$initial_bw <- as.factor(Hmisc::cut2(bw_original$initial_bw, g=10))

############################
df_all <- CreateAllFacet(bw_original, "sex")
a <- ggplot()+
  geom_smooth(data = filter(df_all, !is.na(bodyweight)), 
              aes(x = f_dateorder,
                  y = bodyweight,
                  group=fdose,
                  color = fdose
                  ),
              linetype = 'solid',
              size = 0.5,
              span = 0.6,
              se = FALSE
              ) + facet_wrap(~ facet,dir = 'h') + ggtitle("sex x dose") +     
    scale_colour_manual(values=cc_fdose) + small_axis_theme 
a <- addSmallLegend_8(a)
############################
b <- ggpubr::ggtexttable(subset(dose_pairs, p < 0.3), theme = ggpubr::ttheme("classic")) %>%
 tab_add_title(text = 't-tests; dose pairs < 0.05 asc', padding = unit(1.5, "line")) 

b <- table_cell_bg(b, row = 3:8, column = 2:5,
                      fill="darkolivegreen1")

############################
c <- ggplot()+
  geom_smooth(data =filter(bw_original, !is.na(bodyweight)),
              aes(x = fdose,
                  y = bodyweight,
                  group = interaction(sex,f_dateorder),
                  color = f_dateorder
              ),
              size = 0.5,
              span = 0.6,
              se = FALSE
              ) + facet_wrap(~ sex,dir = 'h') +ggtitle("sex:dateorder interaction")  +
    scale_colour_manual(values=cc_months) + small_axis_theme
c <- addSmallLegend_8(c)
############################
d <- ggplot()+
  geom_smooth(data = filter(bw_original, !is.na(bodyweight) & fdose %in% c(0,10)), 
              aes(x = f_dateorder,
                  y = bodyweight,
                  group = interaction(initial_bw, fdose),
                  color = initial_bw
                  ),
              linetype = 'solid',
              size = 0.5,
              span = 0.6,
              se = FALSE
              ) + ggtitle("sex x dose:initial_bw interaction") +     
    scale_colour_manual(values=cc_initweight) + facet_grid(~ fdose~sex) + small_axis_theme

d <- addSmallLegend_8(d)

pdf("./output/bodyweight/eval_bw_4plot_v1.pdf")
ggpubr::ggarrange(a, b, c, d, common.legend = F, ncol = 2, nrow = 2) 
dev.off()
```


```{r}
 # ggplot()+
 #  geom_smooth(data = filter(bw_original, !is.na(bodyweight)), 
 #              aes(x = factor(date_order),
 #                  y = bodyweight,
 #                  group= interaction(sex,dose),
 #                  color = sex
 #                  ),
 #              linetype = 'solid',
 #              size = 0.4,
 #             # span = 0.6,
 #              alpha = 0.6,
 #              se = FALSE
 #              ) + facet_wrap(~ initial_bw ,dir = 'h', scales = "fixed")+ ggtitle("sex:dose interaction with initial weight bins")
ggplot()+
  geom_smooth(data =filter(bw_original, !is.na(bodyweight)), 
              
              aes(x = fdose,
                  y = bodyweight,
                  group = interaction(sex,f_dateorder),
                  color = date_order,
                  linetype = sex
                  ),
             # linetype = 'solid',
              size = 0.5,
              span = 0.6,
              se = FALSE
              ) + ggtitle("sex:dateorder interaction") + 
  viridis::scale_color_viridis( option = "D", "Month+1")

```

# Variance analysis

## Initial boxplot

```{r}
#https://rpkgs.datanovia.com/ggpubr/reference/ggpar.html
p <- ggpubr::ggboxplot(filter(bw_fit_w_estimates, !is.na(dose)), x = "dose", y = "bodyweight", color='sex')
#ggpubr::ggpar(p, legend = "none")
ggpubr::ggpar(p)
```

## Dose factor

```{r}
anova_fdose <- aov(bodyweight ~ fdose, data = bw_fit_w_estimates)
summary(anova_fdose)
plot(anova_fdose,1)
plot(anova_fdose,2)
broom::tidy(TukeyHSD(anova_fdose, "fdose", conf.level = 0.95))
```

## Dose factor with id

```{r}
anova_id <- aov(bodyweight ~ fdose + sex + sex:fdose + sex:fdose:factor(date_order) , data = bw_fit_w_estimates)
anova_id
anova_id$coefficients

# coefs <- data.frame("id" = gsub("[^0-9.-]", "",names(anova_id$coefficients)), "variance" =gsub("[^0-9.-]", "",anova_id$coefficients)
# ) %>%
#   mutate(variance = as.numeric(variance))%>%
#         inner_join(., bw_fit_w_estimates, by = "id") %>%
#   distinct(id, wave, dose, variance,sex)
# broom::tidy(TukeyHSD(anova_id, "id", conf.level = 0.95))
# plot(anova_id,1)
# plot(anova_id,2)
DT::datatable(broom::tidy(TukeyHSD(anova_id, "date_order", conf.level = 0.95)))
```

## Dose factor with sex and sex:fdose interaction

```{r}
anova_sex <- aov(bodyweight ~ sex + sex:fdose , data = bw_fit_w_estimates)



# is significant on at least one dose
summary(anova_sex)

DT::datatable(broom::tidy(TukeyHSD(anova_sex, "sex:fdose", conf.level = 0.95)))
plot(anova_sex,1)
plot(anova_sex,2)
```

## Dose factor with initial bw

```{r}
# anova_sex <- aov(bodyweight ~ sex + sex:fdose , data = bw_fit_w_estimates)
```